---
title: "R Notebook"
output: html_notebook
---
# Introduction  
A script to remove doublets from the treatment naive subset of GSE123139.  

# Setup  
## Libraries  
```{r}
library("dplyr")
library("ggplot2")
library("Seurat")
library("patchwork")
library("DoubletFinder")
```
## Data  
Data was preprocessed in a separate script and does not include tumour clusters.  
```{r}
GEX = readRDS("./path_to_directory/GSE123139_GEX_treatment_naive.RDS")
cells_metadata = readRDS("./path_to_directory/GSE123139_cells_metadata_treatment_naive.RDS")
```
# Analysis  

```{r}
mel_all = CreateSeuratObject(counts = GEX, meta.data = cells_metadata, project = "mel_t_naive", min.cells = 3, min.features = 200)
mel_all
```
Cleaning memory  
```{r}
rm(GEX)
rm(cells_metadata)
gc()
```
Checking for MT genes
```{r}
mel_all[["percent.mt"]] <- PercentageFeatureSet(mel_all, pattern = "^MT-")
```

```{r}
VlnPlot(mel_all, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```
```{r}
mel_all <- subset(mel_all, subset = nFeature_RNA > 200 & nFeature_RNA < 2500)
mel_all
```

```{r}
mel_all <- NormalizeData(mel_all, normalization.method = "LogNormalize", scale.factor = 10000)
mel_all <- FindVariableFeatures(mel_all, selection.method = "vst", nfeatures = 2000)
```
```{r}
# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(mel_all), 10)

# plot variable features with and without labels
plot1 = VariableFeaturePlot(mel_all)
LabelPoints(plot = plot1, points = top10, repel = TRUE)
```
```{r}
all.genes <- rownames(mel_all)
mel_all <- ScaleData(mel_all, features = all.genes)
```

## PCA  
```{r}
mel_all <- RunPCA(mel_all, features = VariableFeatures(object = mel_all))
```
```{r}
DimPlot(mel_all, reduction = "pca")
```
```{r}
DimPlot(mel_all, reduction = "pca", group.by = "mc_group")
DimPlot(mel_all, reduction = "pca", group.by = "mc_group_granular")
```
```{r}
ElbowPlot(mel_all, ndims = 50)
```

```{r}
mel_all <- FindNeighbors(mel_all, dims = 1:20)
mel_all <- FindClusters(mel_all, resolution = 0.8)
```
## UMAP  
```{r}
mel_all <- RunUMAP(mel_all, dims = 1:20)
```
# Analysis post UMAP  

```{r}
DimPlot(mel_all, reduction = "umap", group.by = "seurat_clusters", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "mc_group", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "mc_group_granular", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "patient_id", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "location", cols = "alphabet")
```
```{r}
FeaturePlot(mel_all, reduction = "umap", features = "nCount_RNA", order = T)
FeaturePlot(mel_all, reduction = "umap", features = "nFeature_RNA", order = T)
```
## Removal of cycling cluster  
Cycling genes from Seurat    
```{r}
s.genes <- cc.genes$s.genes
DotPlot(mel_all, features = s.genes) + coord_flip()
```
Cluster 9 is cycling cells and it is removed  
```{r}
mel_all = mel_all[, mel_all$seurat_clusters !=9]
mel_all
DimPlot(mel_all, reduction = "umap", group.by = "mc_group", cols = "alphabet")
```
# Re-running the pre-processing with new cell composition  
```{r}
mel_all <- FindVariableFeatures(mel_all, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(mel_all)
mel_all <- ScaleData(mel_all, features = all.genes)
mel_all <- RunPCA(mel_all, features = VariableFeatures(object = mel_all))
ElbowPlot(mel_all, ndims = 50)
```
```{r}
mel_all <- FindNeighbors(mel_all, dims = 1:20)
mel_all <- FindClusters(mel_all, resolution = 0.8)
mel_all <- RunUMAP(mel_all, dims = 1:20)
```
```{r}
DimPlot(mel_all, reduction = "umap", group.by = "seurat_clusters", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "mc_group", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "mc_group_granular", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "patient_id", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "location", cols = "alphabet")
```
# Doublet finder  
Source https://github.com/chris-mcginnis-ucsf/DoubletFinder 
```{r}
Sys.time()
sweep.res.list_mel <- paramSweep_v3(mel_all, PCs = 1:20, sct = FALSE)
Sys.time()
```

```{r}
Sys.time()
sweep.stats_mel <- summarizeSweep(sweep.res.list_mel, GT = FALSE)
Sys.time()
```
```{r}
bcmvn_mel <- find.pK(sweep.stats_mel)
```
```{r}
plot(bcmvn_mel$pK, bcmvn_mel$BCmetric)
```
```{r}
bcmvn_mel
```

```{r}
annotations <- mel_all@meta.data$mc_group
homotypic.prop <- modelHomotypic(annotations)
nExp_poi <- round(0.075*nrow(mel_all@meta.data))
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
```
```{r}
Sys.time()
mel_DF <- doubletFinder_v3(mel_all, PCs = 1:20, pN = 0.25, pK = 0.005, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)
Sys.time()
```
```{r}
table(mel_DF$DF.classifications_0.25_0.005_1327)
```
```{r}
DimPlot(mel_DF, reduction = "umap", group.by = "DF.classifications_0.25_0.005_1327")
FeaturePlot(mel_DF, features = "pANN_0.25_0.005_1327", order = TRUE)
```
```{r}
table(mel_DF$DF.classifications_0.25_0.005_1327,
      mel_DF$mc_group)
```
```{r}
VlnPlot(mel_DF, features = "nFeature_RNA", group.by = "DF.classifications_0.25_0.005_1327")
```
```{r}
mel_all_post_treatment_DF = data.frame(mel_DF@meta.data)
head(mel_all_post_treatment_DF)
```
