---
title: "R Notebook"
output: html_notebook
---
# Introduction  
A script analyse B cells from treatment naive samples from GSE123139 after removal of cycling cells and doublets.  

# Setup  
## Libraries  
```{r}
library("dplyr")
library("ggplot2")
library("Seurat")
library("patchwork")
```
## Data  
Data was preprocessed in a separate script and contains only B cell clusters.  
```{r}
metadata_B_cells = read.csv("./path_to_directory/metadata_B_cells.csv", row.names = 1)
head(metadata_B_cells)
```

```{r}
GEX_B_cells = readRDS("./path_to_directory/GEX_B_cells.RDS")
GEX_B_cells[1:5,1:5]
```

```{r}
table(metadata_B_cells$treatment_type)
metadata_B_cells = metadata_B_cells %>% filter(treatment_type == "naive_treatment")
GEX_B_cells = GEX_B_cells[,rownames(metadata_B_cells)]
```

# Analysis  
```{r}
mel_all = CreateSeuratObject(counts = GEX_B_cells, meta.data = metadata_B_cells, project = "mel_B_cells", min.cells = 3, min.features = 200)
mel_all
```
## QC  
```{r}
VlnPlot(mel_all, features = c("nFeature_RNA", "nCount_RNA"))
```
The data was pre-filtered for doublets, so no need to remove anything else.

```{r}
mel_all <- NormalizeData(mel_all, normalization.method = "LogNormalize", scale.factor = 10000)
mel_all <- FindVariableFeatures(mel_all, selection.method = "vst", nfeatures = 500)
```
```{r}
top10 <- head(VariableFeatures(mel_all), 10)

plot1 = VariableFeaturePlot(mel_all)
LabelPoints(plot = plot1, points = top10, repel = TRUE)
```
```{r}
all.genes <- rownames(mel_all)
mel_all <- ScaleData(mel_all, features = all.genes)
```

## PCA  

```{r}
mel_all <- RunPCA(mel_all, features = VariableFeatures(object = mel_all))
```
```{r}
DimPlot(mel_all, reduction = "pca")
```
```{r}
DimPlot(mel_all, reduction = "pca", group.by = "mc_group")
```
```{r}
ElbowPlot(mel_all, ndims = 50)
```

```{r}
mel_all <- FindNeighbors(mel_all, dims = 1:10)
mel_all <- FindClusters(mel_all, resolution = 0.8)
```
## UMAP  
```{r}
mel_all <- RunUMAP(mel_all, dims = 1:10)
```

# Analysis post UMAP  

```{r}
DimPlot(mel_all, reduction = "umap", group.by = "seurat_clusters", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "mc_group", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "mc_group_granular", cols = "alphabet")
DimPlot(mel_all, reduction = "umap", group.by = "patient_id", cols = "alphabet")
```

```{r}
mel_all <- RunTSNE(mel_all, dims = 1:10, verbose = FALSE)
DimPlot(mel_all, reduction = "tsne", group.by = "seurat_clusters", cols = "alphabet")
```

```{r}
FeaturePlot(mel_all, reduction = "umap", features = "nCount_RNA", order = T)
FeaturePlot(mel_all, reduction = "umap", features = "nFeature_RNA", order = T)
```
Re-annotatitng clusters at coarse level for downstream analysis  
```{r}
DotPlot(mel_all, features = c("PTPRC", "CD247", 
                          "CD4", "CD8A", 
                          "NCAM1", "KLRD1", 
                          "CD19", "MS4A1", "CD79A", "CD79B", "CD27", 
                          "CD38", "SDC1", "TNFRSF17",
                          "CR2", "CR1", "FCER2",
                          "LAMP3", "CD83", "ITGAX", 
                          "CD68", "CD163", "ITGAM",
                          "CD14",
                         "FCGR3A"),
        cluster.idents = T,
        dot.scale = 10) + 
  theme_classic() + coord_flip() + ggtitle("Treatment naive")

DotPlot(mel_all, features = c("PTPRC", # CD45
                         "CD2", "CD3D", "CD3E", "CD3G", "CD8A", "CD8B", # T cells
                         "FCGR3A", "NCAM1", "KLRD1", "KIR2DL3", # NK cells
                         "BLNK", "CD19", "MS4A1", "CD79A", "CD79B", "CD27",# B cells
                         "CD38", "SDC1", # plasma cells
                         "CD14", "CD68", "CD163", "CSF1R"), # mono/macrophages
        cluster.idents = T,
        dot.scale = 10) + 
  theme_classic() + coord_flip() + ggtitle("Treatment naive")
```

```{r}
IGH = c("IGHD", "IGHM", "IGHA1", "IGHA2", "IGHG1", "IGHG2", "IGHG3", "IGHG4")
```
```{r}
p1 = VlnPlot(mel_all, features = IGH, ncol=2)
p2 = FeaturePlot(mel_all, features = IGH, ncol=2, order = TRUE)

wrap_plots(p1, p2, ncol = 2)
```
```{r}
p1 = VlnPlot(mel_all, features = IGH, ncol=2)
p2 = FeaturePlot(mel_all, features = IGH, ncol=2, order = TRUE)

wrap_plots(p1, p2, ncol = 2)
```


```{r}
DotPlot(mel_all, features = c("IGHM", "IGHD", "IL4R", "TCL1A", "TXNIP", "FCER2", "FCMR", "SELL", "BANK1", "KLF2",
                              "CD27", "VIM",
                              "CD69", "DUSP1", "JUN", "DNAJB1", "FOS", "HSPA1A",# activated
                              "GPR183", "CD44",
                              "HOPX", "ISG20", "EZR",
                              "IGHG1",  "IGHG2", "IGHG3", "IGHG4", "IGHA1", "IGHA2",
                              "CXCR5",
                          "CD38", "SDC1"
                              ),
        cluster.idents = T,
        dot.scale = 8) + 
  theme_classic() + coord_flip()
DotPlot(mel_all, features = c("IGHM", "IGHD", "IL4R", "TCL1A", "TXNIP", "FCER2", "FCMR", "SELL", "BANK1", "KLF2",
                              "CD27", "VIM",
                              "CD69", "DUSP1", "JUN", "DNAJB1", "FOS", "HSPA1A",# activated
                              "GPR183", "CD44",
                              "HOPX", "ISG20", "EZR",
                              "IGHG1",  "IGHG2", "IGHG3", "IGHG4", "IGHA1", "IGHA2",
                              "CXCR5",
                          "CD38", "SDC1"
                              ),
        cluster.idents = T,
        dot.scale = 8,
        group.by = "cell_type") + 
  theme_classic() + coord_flip()
```

```{r}
mel_all.markers <- FindAllMarkers(mel_all, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
mel_all.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
```
```{r}
mel_all.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(mel_all, features = top10$gene) + NoLegend()
```
