---
title: "CDR3 kidera analysis"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

Analysis of CDR3 kidera characteristics on Silvia's Melanoma data in comparison with the Healthy/Ebola/Covid repertoire data of ours.

```{r setup}
knitr::opts_chunk$set(fig.path = "cdr3_analysis_plots/", dev = "png", dpi = 300)
library(ggplot2)
library(plyr)
```


```{r eval=FALSE}
all_seqs <- readRDS('HV_EB_COV_MEL_comb_CDR_FR_lengths_AA_separated.RDS')

mel <- read.table('Visium_Oct22_slices/results/Visium_all_Aug22_comboDatBoth.txt', sep = '\t', stringsAsFactors = FALSE,
                  header = TRUE)
v_gapped <- c("Visium_Oct22_slices/results/VQuest/2_IMGT-gapped-nt-sequences_visa.txt",
              "Visium_Oct22_slices/results/VQuest/2_IMGT-gapped-nt-sequences_visb.txt")
v_gapped <- lapply(v_gapped, read.table, sep = '\t',
                   stringsAsFactors = FALSE, comment.char = "",
                   header = TRUE, quote = "", fill = TRUE)
v_gapped[[1]] <- v_gapped[[1]][, c("Sequence.ID", "V.REGION")]
v_gapped[[2]] <- v_gapped[[2]][, c("Sequence.ID", "V.REGION")]
v_gapped <- do.call("rbind", v_gapped)
mel <- merge(mel, v_gapped, by.x = "Seq_ID", by.y = "Sequence.ID",
             all.x = TRUE, all.y = FALSE, sort = FALSE)
colnames(mel)[which(colnames(mel) == "V.REGION")] <- "V_gapped"
mel_unique = mel[row.names(unique(mel[,c("Subclass", "V_gapped")])),]
colnames(mel_unique)[which(colnames(mel_unique) == "Patient")] <- "PatientID"
mel_unique$SampleType <- "melanoma"

megafile <- read.table('AntibodyRepertoireCV+RSV+EBOV+YFV.txt', sep = '\t',
                       stringsAsFactors = FALSE, header = TRUE )
megafile <- merge(megafile, all_seqs[, c("Seq_ID", "V_gapped")],
                  by = "Seq_ID", all.x = TRUE, all.y = FALSE)
megafile$BRC_code <- NA
megafile <- megafile[which(megafile$Seq_ID %in% all_seqs$Seq_ID), colnames(mel_unique)]

all_seqs <- rbind(megafile, mel_unique)
saveRDS(all_seqs, 'HV_EB_COV_MEL_comb_Oct22.RDS')

```

```{r}
all_seqs <- readRDS("HV_EB_COV_MEL_comb_Oct22.RDS")
```

# kidera analysis

```{r}
kidera <- all_seqs[, c("Seq_ID", "SampleType", "PatientID", "Subclass", "AA_CDR3_edited", 
               "Boman", "pI_EMBOSS", "Hydrophobicity", 
               colnames(all_seqs)[which(grepl("^kidera", colnames(all_seqs)))])]
```

```{r, dev = "png", dpi=300}
# PCA
kidera <- kidera[which(!is.na(kidera$kidera1.helix.bend.pref)), ]
kidera$SampleType <- replace(kidera$SampleType, which(kidera$SampleType == "melanoma"),
                             "Melanoma")
kidera$SampleType <- factor(kidera$SampleType, 
                            levels = c("Healthy", "Melanoma", "EB", "COVID"),
                            labels = c("HV", "MP", "EB", "Covid"))
kidera <- kidera[which(kidera$Subclass %in% c("IgM", "IgG1", "IgG2", "IgG3", "IgG4", "IgA1", "IgA2", "IgE")), ]
pca <- prcomp( kidera[, colnames(kidera)[which(grepl("^kidera", colnames(kidera)))]])
kidera <- cbind(kidera, pca$x[, 1:2])
ggplot(kidera, aes(x = PC1, y = PC2, colour = SampleType)) + geom_point() + 
  facet_wrap(~ Subclass) + cowplot::theme_cowplot() +
  scale_colour_manual(values = c("HV" = "#d6d6d6", "MP" = "#ff9300", 
                                 "EB" = "#fffb00", "Covid" = "#00f900"), name = "")
```

Too imbalanced the sequence numbers, and all seem to just overlay one another.

Try plot only median for each group and show range as error bars.

```{r}
# modify from ggbiplot::ggbiplot but only get the calculated arrows positions
# representing the PCA loadings
getArrows <- function (pcobj, choices = 1:2, scale = 1, pc.biplot = TRUE, 
          obs.scale = 1 - scale, var.scale = scale, groups = NULL, 
          ellipse = FALSE, ellipse.prob = 0.68, labels = NULL, labels.size = 3, 
          alpha = 1, var.axes = TRUE, circle = FALSE, circle.prob = 0.69, 
          varname.size = 3, varname.adjust = 1.5, varname.abbrev = FALSE) 
{
  library(ggplot2)
  library(plyr)
  library(scales)
  library(grid)
  stopifnot(length(choices) == 2)
  if (inherits(pcobj, "prcomp")) {
    nobs.factor <- sqrt(nrow(pcobj$x) - 1)
    d <- pcobj$sdev
    u <- sweep(pcobj$x, 2, 1/(d * nobs.factor), FUN = "*")
    v <- pcobj$rotation
  }
  else if (inherits(pcobj, "princomp")) {
    nobs.factor <- sqrt(pcobj$n.obs)
    d <- pcobj$sdev
    u <- sweep(pcobj$scores, 2, 1/(d * nobs.factor), FUN = "*")
    v <- pcobj$loadings
  }
  else if (inherits(pcobj, "PCA")) {
    nobs.factor <- sqrt(nrow(pcobj$call$X))
    d <- unlist(sqrt(pcobj$eig)[1])
    u <- sweep(pcobj$ind$coord, 2, 1/(d * nobs.factor), FUN = "*")
    v <- sweep(pcobj$var$coord, 2, sqrt(pcobj$eig[1:ncol(pcobj$var$coord), 
                                                  1]), FUN = "/")
  }
  else if (inherits(pcobj, "lda")) {
    nobs.factor <- sqrt(pcobj$N)
    d <- pcobj$svd
    u <- predict(pcobj)$x/nobs.factor
    v <- pcobj$scaling
    d.total <- sum(d^2)
  }
  else {
    stop("Expected a object of class prcomp, princomp, PCA, or lda")
  }
  choices <- pmin(choices, ncol(u))
  df.u <- as.data.frame(sweep(u[, choices], 2, d[choices]^obs.scale, 
                              FUN = "*"))
  v <- sweep(v, 2, d^var.scale, FUN = "*")
  df.v <- as.data.frame(v[, choices])
  names(df.u) <- c("xvar", "yvar")
  names(df.v) <- names(df.u)
  if (pc.biplot) {
    df.u <- df.u * nobs.factor
  }
  r <- sqrt(qchisq(circle.prob, df = 2)) * prod(colMeans(df.u^2))^(1/4)
  v.scale <- rowSums(v^2)
  df.v <- r * df.v/sqrt(max(v.scale))
  if (obs.scale == 0) {
    u.axis.labs <- paste("standardized PC", choices, sep = "")
  }
  else {
    u.axis.labs <- paste("PC", choices, sep = "")
  }
  u.axis.labs <- paste(u.axis.labs, sprintf("(%0.1f%% explained var.)", 
                                            100 * pcobj$sdev[choices]^2/sum(pcobj$sdev^2)))
  if (!is.null(labels)) {
    df.u$labels <- labels
  }
  if (!is.null(groups)) {
    df.u$groups <- groups
  }
  if (varname.abbrev) {
    df.v$varname <- abbreviate(rownames(v))
  }
  else {
    df.v$varname <- rownames(v)
  }
  df.v$angle <- with(df.v, (180/pi) * atan(yvar/xvar))
  df.v$hjust = with(df.v, (1 - varname.adjust * sign(xvar))/2)
  df.v
}

```

```{r, fig.width=9, fig.height=9}
kidera2 <- list(
  ddply(kidera[which(kidera$SampleType != "MP"), ],
        c("SampleType", "Subclass"), summarise,
                 PC1_median = median(PC1),
                 PC1_lowerq = quantile(PC1, 0.025),
                 PC1_upperq = quantile(PC1, 0.975),
                 PC2_median = median(PC2),
                 PC2_lowerq = quantile(PC2, 0.025),
                 PC2_upperq = quantile(PC2, 0.975)),
  ddply(kidera[which(kidera$SampleType == "MP"), ], 
        c("SampleType", "PatientID", "Subclass"), summarise,
                 PC1_median = median(PC1),
                 PC1_lowerq = quantile(PC1, 0.025),
                 PC1_upperq = quantile(PC1, 0.975),
                 PC2_median = median(PC2),
                 PC2_lowerq = quantile(PC2, 0.025),
                 PC2_upperq = quantile(PC2, 0.975))
)
kidera2[[1]]$PatientID <- NA
kidera2[[1]] <- kidera2[[1]][, colnames(kidera2[[2]])]
kidera2 <- do.call("rbind", kidera2)

arrows <- getArrows( pca )
arrows$varname <- gsub("\\..*", "", arrows$varname)
arrows$varname <- gsub("kidera", "K", arrows$varname)

# Plot
ggplot(kidera2, aes(x = PC1_median, y = PC2_median, colour = SampleType)) +
    scale_colour_manual(values = c("HV" = "#d6d6d6", "MP" = "#ff9300", 
                                   "EB" = "#fffb00", "Covid" = "#00f900"), name = "") +
    geom_point() + geom_errorbar(aes(xmin = PC1_lowerq, xmax = PC1_upperq, alpha = (SampleType == "MP"))) +
    geom_errorbar(aes(ymin = PC2_lowerq, ymax = PC2_upperq, alpha = (SampleType == "MP"))) +
    ggrepel::geom_text_repel(aes(label = PatientID)) + xlab("PC1") + ylab("PC2") + 
  scale_alpha_manual(values = c("FALSE" = 1, "TRUE" = 0.3), guide = "none") +
    cowplot::theme_cowplot() + facet_wrap(~ Subclass)
```

```{r, fig.width=9, fig.height=3.5}
kidera2$Subclass <- factor(kidera2$Subclass,
                           levels = c("IgM", "IgG1", "IgG2", "IgG3", "IgG4", "IgA1", 
                                      "IgA2", "IgE"))

cowplot::plot_grid(
  ggplot(kidera2[which(kidera2$Subclass %in% c("IgA1", "IgG1", "IgG2", "IgG3", "IgM")), ], 
       aes(x = PC1_median, y = PC2_median, colour = SampleType)) +
    scale_colour_manual(values = c("HV" = "#d6d6d6", "MP" = "#ff9300", 
                                   "EB" = "#fffb00", "Covid" = "#00f900"), name = "") +
    geom_point() + geom_errorbar(aes(xmin = PC1_lowerq, xmax = PC1_upperq, alpha = (SampleType == "MP"))) +
    geom_errorbar(aes(ymin = PC2_lowerq, ymax = PC2_upperq, alpha = (SampleType == "MP"))) +
    ggrepel::geom_text_repel(aes(label = PatientID)) + xlab("PC1") + ylab("PC2") + 
  scale_alpha_manual(values = c("FALSE" = 1, "TRUE" = 0.3), guide = "none") +
    cowplot::theme_cowplot() + facet_wrap(~ Subclass),
  
  cowplot::plot_grid(
    NULL,
    ggplot(arrows) + geom_segment(aes(x = -0.5, y = 0.5, 
                                xend = xvar - 0.5, yend = yvar + 0.5),
             arrow = arrow(length = unit(0.5, "picas")), color = "grey70") +
    ggrepel::geom_text_repel(aes(x = xvar - 0.5, y = yvar + 0.5, label = varname), 
              color = "grey70", size = 3) + cowplot::theme_cowplot() + xlab("PC1") + ylab("PC2"),
    nrow = 2, ncol = 1
  ),
  
  ncol = 2, rel_widths = c(3, 1)

)

```

all patients represented individually

```{r, fig.width=10, fig.height=5.5}
# all patients represented individually
kidera4 <- ddply(kidera, 
        c("SampleType", "PatientID", "Subclass"), summarise,
                 PC1_median = median(PC1),
                 PC1_lowerq = quantile(PC1, 0.025),
                 PC1_upperq = quantile(PC1, 0.975),
                 PC2_median = median(PC2),
                 PC2_lowerq = quantile(PC2, 0.025),
                 PC2_upperq = quantile(PC2, 0.975))

ggplot(kidera4, 
       aes(x = PC1_median, y = PC2_median, colour = SampleType)) +
    scale_colour_manual(values = c("HV" = "#d6d6d6", "MP" = "#ff9300", 
                                   "EB" = "#fffb00", "Covid" = "#00f900"), name = "") +
    geom_point() + geom_errorbar(aes(xmin = PC1_lowerq, xmax = PC1_upperq), alpha = 0.3) +
    geom_errorbar(aes(ymin = PC2_lowerq, ymax = PC2_upperq), alpha = 0.3) +
    xlab("PC1") + ylab("PC2") + 
    cowplot::theme_cowplot() + facet_grid(SampleType ~ Subclass)

```

```{r, fig.width=8, fig.height=3}
kidera3 <- reshape2::melt(kidera, id.vars = c("SampleType", "PatientID", "Subclass"),
                          measure.vars = colnames(kidera)[which(grepl("kidera", colnames(kidera)))]) 
kidera3$Subclass <- factor(kidera3$Subclass,
                           levels = c("IgM", "IgG1", "IgG2", "IgG3", "IgG4", "IgA1", 
                                      "IgA2", "IgE"))
kidera3$variable <- factor(kidera3$variable, levels = levels(kidera3$variable),
                           labels = c("K1 (helix/bend)", "K2 (side-chain size)",
                                      "K3 (extended structure)", "K4 (hydrophobicity)",
                                      "K5 (double-bend)", "K6 (partial specific volume)",
                                      "K7 (flat extended)", "K8 (alpha region)",
                                      "K9 (pK-C)", "K10 (surrounding hydrophobicity)"))

kidera3$type <- kidera3$PatientID
kidera3$type <- replace(kidera3$type, 
                        which(kidera3$SampleType == "HV"),
                        "HV")

test2 <- ddply(kidera3[kidera3$SampleType %in% c("MP", "HV"), ],
        c("SampleType", "type", "Subclass", "variable"),
        summarise,
        median = median(value))

test <- ggpubr::compare_means(
  value ~ type, 
  data = kidera3[which(kidera3$SampleType %in% c("HV", "MP")), ], 
  group.by = c("variable", "Subclass")
)

test$effsize <- apply(test, MARGIN =1, function(x){
  effsize::cliff.delta(
    kidera3[kidera3$variable == x[1] & kidera3$Subclass == x[2] &
               kidera3$type == x[4], "value"], 
    kidera3[kidera3$variable == x[1] & kidera3$Subclass == x[2] &
               kidera3$type == x[5], "value"])$estimate
})

test$group1_median <- apply(test, MARGIN =1, function(x){
  test2[test2$type == x[4] & test2$Subclass == x[2] &
          test2$variable == x[1], "median"]
})
test$group2_median <- apply(test, MARGIN =1, function(x){
  test2[test2$type == x[5] & test2$Subclass == x[2] &
          test2$variable == x[1], "median"]
})
test$diff <- test$group2_median - test$group1_median
```

```{r, fig.width=9, fig.height=9}
test$variable <- factor(test$variable, levels = levels(test$variable),
                           labels = c("K1\n(helix/bend)", "K2\n(side-chain\nsize)",
                                      "K3\n(extended\nstructure)", "K4\n(hydrophobicity)",
                                      "K5\n(double-bend)", "K6 (partial\nspecific volume)",
                                      "K7 (flat\nextended)", "K8 (alpha\nregion)",
                                      "K9 (pK-C)", "K10\n(surrounding\nhydrophobicity)"))

test <- test[which(test$Subclass %in% c("IgM", "IgG1", "IgG2",
                                        "IgG3", "IgA1")), ]
test$p.adj <- p.adjust(test$p)
test$signif <- -log10(test$p.adj)
test$signif <- ifelse(test$signif >= 5e-10, -log10(5e-10), test$signif)
test$signif <- (test$p.adj < 0.05)

ggplot(test[test$group1 == "HV", ], 
       aes(x = group2, y = Subclass, 
           color = diff, size = signif)) +
  geom_point() + cowplot::theme_cowplot() +
  facet_wrap(~ variable) +
  scale_color_gradient2(low = "blue", high = "red",
                        name = "difference in\nCDR3 kidera\nvalue vs HV") + xlab("") + ylab("") +
  scale_size_manual(values = c("TRUE" = 6, "FALSE" = 1), 
                    name = "adjusted p<0.05") +
#  scale_size_continuous(name = "adjusted\np-value",
#                        breaks = c(-log10(1), -log10(0.5),
#                                   -log10(0.05), -log(0.005)), 
#                        labels = c("1", "0.5", "0.05", "0.005")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

only selected Kidera factors

```{r, fig.width=7, fig.height=3}
ggplot(test[test$group1 == "HV" &
              grepl("K1\n|K3|K8", test$variable), ], 
       aes(x = group2, y = Subclass, 
           color = diff, size = signif)) +
  geom_point() + cowplot::theme_cowplot() +
  facet_wrap(~ variable) +
  scale_color_gradient2(low = "blue", high = "red",
                        name = "difference in\nCDR3 kidera\nvalue vs HV") + xlab("") + ylab("") +
  scale_size_manual(values = c("TRUE" = 6, "FALSE" = 1), 
                    name = "adjusted p<0.05") +
#  scale_size_continuous(name = "adjusted\np-value",
#                        breaks = c(-log10(1), -log10(0.5),
#                                   -log10(0.05), -log(0.005)), 
#                        labels = c("1", "0.5", "0.05", "0.005")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
write.csv(test, 'cdr3_kidera_stats.csv', row.names = FALSE)
```

# Amino acid usage & secondary structures

```{r}
# Average flexibility index from Bhaskaran & Ponnuswamy
# avialable in ProtScale in ExPASy
# https://web.expasy.org/protscale/pscale/Averageflexibility.html

flexibility_scale <- c("A" = 0.360, "R" = 0.530, "N" = 0.460, "D" = 0.510, 
                       "C" = 0.350, "Q" = 0.490, "E" = 0.500, "G" = 0.540,  
                       "H" = 0.320, "I" = 0.460, "L" = 0.370, "K" = 0.470,  
                       "M" = 0.300, "F" = 0.310, "P" = 0.510, "S" = 0.510,  
                       "T" = 0.440, "W" = 0.310, "Y" = 0.420, "V" = 0.390)  

all_seqs$flexibility <- apply(all_seqs[, c("Num_AAs", Biostrings::AA_STANDARD)], MARGIN = 1, function(x){
  sum(x[2:21] * flexibility_scale) / x[1]
})
```

```{r, fig.width=8, fig.height=4}
aacount <- all_seqs[, c("Seq_ID", "PatientID", "SampleType", "UseAsRef",
                        "Subclass", "Num_AAs", "flexibility",
                        Biostrings::AA_STANDARD)]
aacount[, c(Biostrings::AA_STANDARD)] <- aacount[, c(Biostrings::AA_STANDARD)] / aacount$Num_AAs

# count % residues forming helix/strand/coil
aacount$helix <- apply(aacount[, c("E", "A", "L", "M", "Q", "K", "R", "H")], MARGIN = 1,
                       sum)
aacount$strand <- apply(aacount[, c("V", "I", "Y", "C", "W", "F", "T")], MARGIN = 1,
                        sum)
aacount$coil <- apply(aacount[, c("G", "N", "P", "S", "D")], MARGIN = 1,
                      sum)
aacount <- aacount[which(aacount$SampleType %in% c("Healthy", "melanoma")), ]
aacount$SampleType <- factor(aacount$SampleType, levels = c("Healthy", "melanoma"), labels = c("HV", "MP"))
aacount$type <- aacount$PatientID
aacount$type <- replace(aacount$type, which(aacount$SampleType == "HV"), "HV")

aacount2 <- reshape2::melt(aacount, id.vars = c("Seq_ID", "UseAsRef", "type", "Subclass"), 
                           measure.vars = Biostrings::AA_STANDARD)
aacount3 <- reshape2::melt(aacount, id.vars = c("Seq_ID", "UseAsRef", "type", "Subclass"), 
                           measure.vars = c("helix", "strand", "coil", "flexibility"))

ggplot(ddply(aacount2[which(aacount2$Subclass %in% c("IgM", "IgG1", "IgG2", "IgG3", "IgA1")), ], 
             c("type", "Subclass", "variable"), 
             summarise, median = median(value, na.rm = TRUE)), 
       aes(y = type, x = variable, fill = median)) + geom_tile() +
  scale_fill_gradient2(na.value = "white", name = "CDR3 AA\nusage\n(median %)", 
                       labels = scales::percent) + 
  cowplot::theme_cowplot() + facet_wrap(~ Subclass) + xlab("Amino acids") + ylab("")
```

```{r, fig.width=7, fig.height=3}
test <- ggpubr::compare_means(value ~ type, 
                              data = aacount3[aacount3$Subclass %in% c("IgM", "IgA1", "IgG1", "IgG2", "IgG3"), ], group.by = c("variable", "Subclass"))
test$effsize <- apply(test, MARGIN =1, function(x){
  effsize::cliff.delta(
    aacount3[aacount3$variable == x[1] & aacount3$Subclass == x[2] &
               aacount3$type == x[4], "value"], 
    aacount3[aacount3$variable == x[1] & aacount3$Subclass == x[2] &
               aacount3$type == x[5], "value"])$estimate
})

aacount4 <- ddply(aacount3[which(aacount3$Subclass %in% c("IgM", "IgG1", "IgG2", "IgG3", "IgA1")), ], 
             c("type", "Subclass", "variable"), 
             summarise, median = median(value, na.rm = TRUE))

test$group1_median <- apply(test, MARGIN = 1, function(x){
  aacount4[which(aacount4$type == x[4] & 
                   aacount4$variable == x[1] &
                   aacount4$Subclass == x[2]), "median"]
})

test$group2_median <- apply(test, MARGIN = 1, function(x){
  aacount4[which(aacount4$type == x[5] & 
                   aacount4$variable == x[1] &
                   aacount4$Subclass == x[2]), "median"]
})

test$diff <- test$group2_median - test$group1_median
test$p.adj <- p.adjust(test$p)
test$signif <- -log10(test$p.adj)
test$signif <- ifelse(test$signif >= 5e-10, -log10(5e-10), test$signif)
test$signif <- (test$p.adj < 0.05)
test <- test[which(test$group1 == "HV"), ]

ggplot(test[which(test$variable != "flexibility"), ], 
       aes(x = group2, y = Subclass, 
           color = diff, size = signif)) +
  geom_point() + cowplot::theme_cowplot() +
  facet_wrap(~ variable) +
  scale_color_gradient2(low = "blue", high = "red",
                        name = "difference in\n% CDR residue\nusage vs HV", labels = scales::percent) + 
  xlab("") + ylab("") +
  scale_size_manual(values = c("TRUE" = 6, "FALSE" = 1), 
                    name = "adjusted p<0.05") +
#  scale_size_continuous(name = "adjusted\np-value",
#                        breaks = c(-log10(1), -log10(0.5),
#                                   -log10(0.05), -log(0.005)), 
#                        labels = c("1", "0.5", "0.05", "0.005")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


```{r, fig.width=3.5, fig.height=3}
ggplot(test[which(test$variable == "flexibility"), ], 
       aes(x = group2, y = Subclass, 
           color = diff, size = signif)) +
  geom_point() + cowplot::theme_cowplot() +
  facet_wrap(~ variable) +
  scale_color_gradient2(low = "blue", high = "red") + 
  xlab("") + ylab("") +
  scale_size_manual(values = c("TRUE" = 6, "FALSE" = 1), 
                    name = "adjusted p<0.05") +
#  scale_size_continuous(name = "adjusted\np-value",
#                        breaks = c(-log10(1), -log10(0.5),
#                                   -log10(0.05), -log(0.005)), 
#                        labels = c("1", "0.5", "0.05", "0.005")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
write.csv(test, 'cdr3_SS_stats.csv', row.names = FALSE)
```

# collapse one sequence per clone

```{r}
test <- ggpubr::compare_means(value ~ type, 
                              data = aacount3[which(aacount3$Subclass %in% c("IgM", "IgA1", "IgG1", "IgG2", "IgG3") & aacount3$UseAsRef), ], group.by = c("variable", "Subclass"))
test$effsize <- apply(test, MARGIN =1, function(x){
  y1 <- aacount3[which(aacount3$variable == x[1] & aacount3$Subclass == x[2] &
               aacount3$type == x[4] & aacount3$UseAsRef), "value"]
  y2 <- aacount3[which(aacount3$variable == x[1] & aacount3$Subclass == x[2] &
               aacount3$type == x[5] & aacount3$UseAsRef), "value"]
  if( length(y1) > 3 & length(y2) > 3 ){
    return(effsize::cliff.delta(y1, y2)$estimate)
  } else return(NA)
})

aacount4 <- ddply(aacount3[which(aacount3$Subclass %in% c("IgM", "IgG1", "IgG2", "IgG3", "IgA1") & aacount3$UseAsRef), ], 
             c("type", "Subclass", "variable"), 
             summarise, median = median(value, na.rm = TRUE))

test$group1_median <- apply(test, MARGIN = 1, function(x){
  aacount4[which(aacount4$type == x[4] & 
                   aacount4$variable == x[1] &
                   aacount4$Subclass == x[2]), "median"]
})

test$group2_median <- apply(test, MARGIN = 1, function(x){
  aacount4[which(aacount4$type == x[5] & 
                   aacount4$variable == x[1] &
                   aacount4$Subclass == x[2]), "median"]
})

#test$diff <- test$group2_median - test$group1_median
test$diff <- apply(test[, c("group1_median", "group2_median")], MARGIN = 1, function(x){
  o <- try(x[2] - x[1], silent = TRUE)
  if( class(o) != "try-error") return(o) else return(NA)
})
test$p.adj <- p.adjust(test$p)
test$signif <- -log10(test$p.adj)
test$signif <- ifelse(test$signif >= 5e-10, -log10(5e-10), test$signif)
test$signif <- (test$p.adj < 0.05)
test <- test[which(test$group1 == "HV"), ]
```

```{r}
saveRDS(test, 'cdr3_SS_stats_UseAsRefTrue.rds')
```

```{r, fig.width=7, fig.height=3}
ggplot(test[which(test$variable != "flexibility"), ], 
       aes(x = group2, y = Subclass, 
           color = diff, size = signif)) +
  geom_point() + cowplot::theme_cowplot() +
  facet_wrap(~ variable) +
  scale_color_gradient2(low = "blue", high = "red",
                        name = "difference in\n% CDR residue\nusage vs HV", labels = scales::percent) + 
  xlab("") + ylab("") +
  scale_size_manual(values = c("TRUE" = 6, "FALSE" = 1), 
                    name = "adjusted p<0.05") +
#  scale_size_continuous(name = "adjusted\np-value",
#                        breaks = c(-log10(1), -log10(0.5),
#                                   -log10(0.05), -log(0.005)), 
#                        labels = c("1", "0.5", "0.05", "0.005")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r, fig.width=3.5, fig.height=3}
ggplot(test[which(test$variable == "flexibility"), ], 
       aes(x = group2, y = Subclass, 
           color = diff, size = signif)) +
  geom_point() + cowplot::theme_cowplot() +
  facet_wrap(~ variable) +
  scale_color_gradient2(low = "blue", high = "red") + 
  xlab("") + ylab("") +
  scale_size_manual(values = c("TRUE" = 6, "FALSE" = 1), 
                    name = "adjusted p<0.05") +
#  scale_size_continuous(name = "adjusted\np-value",
#                        breaks = c(-log10(1), -log10(0.5),
#                                   -log10(0.05), -log(0.005)), 
#                        labels = c("1", "0.5", "0.05", "0.005")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r}
sessionInfo()
```